### 2014-02-05 docs/thesis/chap-related_work.tex

\chapter{Related Work}
\label{chapter-related_work}
\section{Status of MIR}
\subsection{TREC}
\subsection{TREC}
\subsection{CLEF}
\subsection{NTCIR}
\section{Mathematical based search projects}
\subsection{MIAS}
\subsection{EgoMath}
\subsection{DLMFSearch}
\subsection{MathWebSearch}
\section{Other related projects and initiatives}
\subsection{arXMLiv}
\thispagestyle{empty}