### 2014-02-05 docs/thesis/chap-technical_considerations.tex

\chapter{Technical Considerations}
\label{chapter-technical_considerations}
In the previos chapter we presented different abstract models which can be suitable in the building of a math based search system.
In this section, we present different aspects that need to be considered when mapping from these theoretical models to a concrete implementation.

\section{Digital Representations of Mathematics}
\subsection{MathML}
\subsection{OpenMath}
\subsection{LaTeX}

\section{Math Extraction Tools}
\section{From LaTeX to MathML}
\subsection{From PDF to MathML}
\subsubsection{Maxtract}
\subsubsection{InftyReader}
\subsubsection{OCR analysis}
\subsection{Other projects}

\section{Computational Algebraic Systems}
In the previous section we discussed about algorithmical ways to tranform, simplify and detect patterns in mathematical equations. Here, we explore what tools are available for performing this type of operations and how they are suitable for our system.

\subsection{Mathematica}
\subsection{Mapple}
\subsection{Matlab}
\subsection{Octane}
\subsection{SAGE}