### 2014-02-05 docs/thesis/chap-theoretical_background.tex

\chapter{Theoretical Background}
\label{chapter-theoretical_background}
Information Retrieval (IR) consists on the activity of identifying relevant documents from a collection based on some input parameters.
Different models have been developed through the evolution of the field. In the subsequent sections, the most relevant to this work are presented.

\section{Boolean Model}
This model was the first one to be developed and is currently one of the most used in the current implementation of IR systems because of its simplicity.

\section{Vector Space Model}

\thispagestyle{empty}