### 2014-02-21 docs/thesis/chap-conclusions.tex
added section of dlmf

\chapter{Conclusions}
\label{chapter-conclusions}