### 2014-02-21 docs/thesis/chap-introduction.tex

\chapter{Introduction}

\section{Context}

\subsection{Invenio}
\subsection{CDS}

\section{Goals}