### 2014-02-24 docs/thesis/appendix-uml.tex

\chapter{Detailed system architecture diagrams}

\section{Analyzers}
\begin{figure}[h!]
	\centering
	\includegraphics[height=4 cm]{visio_drawings/analyzers.png}
	\caption{Overview of {\small{\codefont cern.ch.mathexplorer.lucene.analysis.analyzers}} package}
	\label{uml_analyzers}
\end{figure}

\section{Tokenizers}
\begin{figure}[h!]
	\centering
	\includegraphics[height=4 cm]{visio_drawings/tokenizers.png}
	\caption{Overview of {\small{\codefont cern.ch.mathexplorer.lucene.analysis.tokenizers}} package}
	\label{uml_tokenizers}
\end{figure}

\section{Filters}
\begin{figure}[h!]
	\centering
	\includegraphics[height=4 cm]{visio_drawings/filters.png}
	\caption{Overview of {\small{\codefont cern.ch.mathexplorer.lucene.analysis.filters}} package}
	\label{uml_filters}
\end{figure}

\section{Query Parsers}
\begin{figure}[h!]
	\centering
	\includegraphics[height=4 cm]{visio_drawings/query.png}
	\caption{Overview of {\small{\codefont cern.ch.mathexplorer.lucene.query}} package}
	\label{uml_filters}
\end{figure}

\section{Packages Relations}
\begin{figure}[h!]
	\centering
	\includegraphics[height=12 cm]{visio_drawings/main_packages.png}
	\caption{Overview of the relations between the main packages}
	\label{uml_filters}
\end{figure}