### 2014-02-24 docs/thesis/chap-future_work.tex

\chapter{Future work}
\label{chapter-future_work}