### 2014-02-25 docs/thesis/appendix-test_equations.tex
work on thesis: finished sample equations

\chapter{Test equations}

\begin{longtable}{|m{0.75cm}|>
{\centering\arraybackslash}m{2.7cm}|>
{\centering\arraybackslash}m{6.5cm}|>
{\centering\arraybackslash}m{5cm}|
}
\hline 
\textbf{Id} & 
\textbf{Description} &
\textbf{MathML code} &
\textbf{Visual Form}

\\
\hline
Eq1 & \scriptsize{ Time-independent Schr\char"00F6dinger equation} & \scriptsize {\codefont <math><mfenced close=")" open="("> <mrow> <mo>-</mo> <mfrac> <msup> <mi>ℏ</mi> <mn>2</mn> </msup> <mrow> <mn>2</mn> <msub> <mi>m</mi> <mi>e</mi> </msub> </mrow> </mfrac> <msup> <mo>$\nabla$</mo> <mn>2</mn> </msup> <mo>+</mo> <mi>V</mi> <mfenced close=")" open="("> <mi>r</mi> </mfenced> </mrow> </mfenced> <mi>ψ</mi> <mfenced close=")" open="("> <mi>r</mi> </mfenced> <mo>=</mo> <mi>E</mi> <mi>ψ</mi> <mfenced close=")" open="("> <mi>r</mi> </mfenced> </math>} & \small{ $\left (-\frac{\hbar^{2}}{2m_e}\nabla^{2} + V(\mathbf{r})\right )\psi(\mathbf{r}) = E\psi(\mathbf{r})$ }\\ \hline

Eq2 & \scriptsize{Atomic mass in terms of number of proton and number of neutrons} & \scriptsize {\codefont <math> <mi>A</mi> <mo>=</mo> <mi>Z</mi> <mo>+</mo> <mi>N</mi> </math>} & \small{$A = Z + N$} \\ \hline

Eq3 & \scriptsize{Radiation flux in term of the initial intensity, the absorption coefficient and the thickness of a substance} & \scriptsize {\codefont <math> <mi>I</mi> <mo>=</mo> <msub> <mi>I</mi> <mn>0</mn> </msub><msup> <mi>e</mi> <mrow> <mo>-</mo> <mi>μ</mi> <mi>x</mi> </mrow> </msup> </math>
} & \small{$I = I_0e^{-\mu x}$ } \\ \hline

Eq4 & \scriptsize{Higgs Boson decaying process into $W^+$ and $W^-$ bosons. This is not exactly a mathematical equation, but a physical process. However it is also expressed in LaTeX and to high energy physics experts, it makes sense to search them.} & \scriptsize {\codefont <math> <mi>H</mi> <mo>→</mo> <msup> <mi>W</mi> <mo>+</mo> </msup> <msup> <mi>W</mi> <mo>-</mo> </msup> </math>} & \small{$H\rightarrow W^+W^-$ } \\ \hline

Eq5 & \scriptsize{Einstein Field equation} & \scriptsize {\codefont <math> <msub> <mi>G</mi> <mrow> <mi>μ</mi> <mi>ν</mi> </mrow> </msub> <mo>=</mo> <msub> <mi>R</mi> <mrow> <mi>μ</mi> <mi>ν</mi> </mrow> </msub> <mo>-</mo> <mfrac> <mn>1</mn> <mn>2</mn> </mfrac> <mi>R</mi> <msub> <mi>g</mi> <mrow> <mi>μ</mi> <mi>ν</mi> </mrow> </msub> <mo>=</mo> <mfrac> <mrow> <mn>8</mn> <mi>π</mi> <mi>G</mi> </mrow> <msup> <mi>c</mi> <mn>4</mn> </msup> </mfrac> <msub> <mi>T</mi> <mrow> <mi>μ</mi> <mi>ν</mi> </mrow> </msub> </math>
 </math>
 } & \footnotesize{$G_{\mu\nu} = R_{\mu\nu} - {\textstyle 1 \over 2}R\,g_{\mu\nu} = {8 \pi G \over c^4} T_{\mu\nu}$ } \\ \hline

Eq6 & \scriptsize{Luminosity of a star in terms of its temperature and radius} & \scriptsize {\codefont <math> <mi>L</mi> <mo>=</mo> <mn>4</mn> <mi>π</mi> <msup> <mi>R</mi> <mn>2</mn> </msup> <mi>σ</mi> <msubsup> <mi>T</mi> <mrow> <mi>e</mi> <mi>f</mi> <mi>f</mi> </mrow> <mn>4</mn> </msubsup> </math>
} & \small{$ L = 4 \pi R^2 \sigma T^4_{eff} $ } \\ \hline

Eq7 & \scriptsize{Free-cyclotron motion of an electron in terms of its charge, mass and the magnetic field} & \scriptsize {\codefont <math> <msub> <mi>ω</mi> <mi>c</mi> </msub> <mo>=</mo> <mo>|</mo> <mi>q</mi> <mover accent="true"> <mi>B</mi> <mo>→</mo> </mover> <mo>|</mo> <mo>/</mo> <mi>m</mi> </math>
} & \small{$ \omega_c = |q\vec{B}|/m $ } \\ \hline

Eq8 & \scriptsize{Spin energy of a particle in presence of an external magnetic field} & \scriptsize{\codefont <math> <mi>E</mi> <mo>=</mo> <mo>-</mo> <mover accent="true"> <mi>μ</mi> <mo>→</mo> </mover> <mo>⋅</mo> <mover accent="true"> <mi>B</mi> <mo>→</mo> </mover> </math>
} & \small { $E = -\vec{\mu} \cdot \vec{B}$ } \\ \hline

Eq9 & \scriptsize{Quantum electrodynamics Lagrangian modelling the interaction of the -1/2 spin field with the electromagnetic field} & \scriptsize{\codefont <math > <msub> <mi mathvariant="script">L</mi> <mstyle mathvariant="normal"> <mi>Q</mi> <mi>E</mi> <mi>D</mi> </mstyle> </msub> <mo>=</mo> <msub> <mover accent="true"> <mi>ψ</mi> <mo>̄</mo> </mover> <mi>i</mi> </msub> <mi>i</mi> <mfenced close=")" open="("> <mrow> <msub> <mfenced close=")" open="("> <mrow> <msup> <mi>γ</mi> <mi>μ</mi> </msup> <msub> <mi>D</mi> <mi>μ</mi> </msub> </mrow> </mfenced> <mrow> <mi>i</mi> <mi>j</mi> </mrow> </msub> <mo>-</mo> <mi>m</mi> <msub> <mi>δ</mi> <mrow> <mi>i</mi> <mi>j</mi> </mrow> </msub> </mrow> </mfenced> <msub> <mi>ψ</mi> <mi>j</mi> </msub> <mo>-</mo> <mfrac> <mn>1</mn> <mn>4</mn> </mfrac> <msub> <mi>F</mi> <mrow> <mi>μ</mi> <mi>ν</mi> </mrow> </msub> <msup> <mi>F</mi> <mrow> <mi>μ</mi> <mi>ν</mi> </mrow> </msup> </math>
} & \small { $\mathcal{L}_\mathrm{QED} 
= \bar{\psi}_i i \left( (\gamma^\mu D_\mu)_{ij} - m\, \delta_{ij}\right) \psi_j - \frac{1}{4}F_{\mu \nu} F^{\mu \nu}  $ } \\ \hline

Eq10 & \scriptsize{Einstein-Hilbert action} & \scriptsize{\codefont <math> <mi>S</mi> <mo>=</mo> <mfrac> <mn>1</mn> <mrow> <mn>2</mn> <mi>κ</mi> </mrow> </mfrac> <mo>∫</mo> <mi>R</mi> <msqrt> <mrow> <mo>-</mo> <mi>g</mi> </mrow> </msqrt> <msup> <mi>d</mi> <mn>4</mn> </msup> <mi>x</mi> </math>} & \small { $ 
    S = {1 \over 2\kappa} \int R \sqrt{-g} \, \mathrm{d}^4x	$ } \\ \hline

\caption{Sample equations used during testing}
\label{equations_table}
\end{longtable}