### 2014-02-28 docs/thesis/chap-advanced_techniques.tex
added screenshots

\chapter{Advanced Techniques}
\label{chap-advances_techniques}

\section{Latent Semantic Indexing with Semantic Vectors}

\section{Latent Dirichlet Allocation with Apache Mahout}