### 2014-03-05 docs/thesis/chap-future_work.tex

\chapter{Future work}
\label{chap-future_work}

\subsection{Features extraction}

\subsection{Advanced Techniques}

\subsection{End to end system}