### 2014-11-23 jcdl2015/paper/main.tex
Completed v1 of demo version of paper.

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%Imports%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[mathletters]{ucs}
\usepackage{amssymb}
\usepackage[utf8x]{inputenc}
\usepackage[cyr]{aeguill}
\usepackage[english]{babel}
\usepackage{fancyhdr, ifpdf}
\usepackage{array}
\usepackage{multirow}
\usepackage{textcomp}
\usepackage{verbatim} 
\usepackage{listings}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage{fontspec}
\usepackage{booktabs}
\usepackage{amsmath,amssymb,amstext,mathtext}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{url}
\usepackage{spverbatim}
\usepackage{changepage}
\usepackage{caption}
\usepackage{subcaption}

\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother

\newcommand{\loadencoding}[1]{%
  \begingroup
  \def\DeclareUnicodeCharacter##1##2{%
    \expandafter\gdef\csname uni\number"##1\endcsname
      {{\fontencoding{#1}\selectfont##2}}}
  \makeatletter
  \lowercase{\input{#1enc.dfu}}%
  \endgroup}
\newcommand\unisymbol[1]{\csname uni\number#1\endcsname}
\loadencoding{T1}
\loadencoding{T2A}
\newfontfamily\unicodefont{Arial Unicode MS}
\newfontfamily\codefont{Courier}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{JCDL}{2014 Knoxville, USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{$5e^{x+y}$: A Math Aware Search Engine (for CDS)}

%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\author{
\alignauthor Arthur Oviedo\\
  \affaddr{EPFL, Switzerland.}\\%
  \affaddr{arthur.oviedo@alumini.epfl.ch}%
\alignauthor Nikolaos Kasioumis \\
  \affaddr{CERN, Switzerland.}\\%
  \affaddr{nikos.kasioumis@cern.ch}%
\alignauthor Karl Aberer\\%
  \affaddr{EPFL, Switzerland.}\\%
  \affaddr{karl.aberer@epfl.ch}%
}

\maketitle
\begin{abstract}
This paper presents $5e^{x+y}$, a system that complements CERN Document Server (CDS), by adding extracting, indexing and querying of mathematical content, expressed as mathematical equations.
\end{abstract}

% A category with the (minimum) three required fields
\category{X.X}{Digital Libraries}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{Y.Y}{Information Retrieval}{Metrics}

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
CDS\cite{CDS1} (CERN Document Server) is the institutional digital library system
developed and used at CERN\cite{CERN}. It contains more than 1 million records and stores more than 400000 full text documents, organized in different collections among published articles, books, preprints. Invenio\cite{invenio} is the open-source digital library software platform behind CDS and is developed in parallel to CDS at CERN. Besides CDS, Invenio supports around thirty scientific institutions worldwide including INSPIRE (a collaboration between Fermilab, CERN, DESY and SLAC) EPFL and ILO. 
An important part of the CDS content contains mathematical content, represented in the form of equations. Only the Preprints collection contains more than 700000 records, harvesting documents from services like ArXiV\cite{arxiv} where most of the documents are in the areas of physics, mathematics and statistics which are very rich in this type of content. Even though Invenio provides a powerful searching mechanism that allows to query the stored records by different fields like author or keywords, these approaches are not suitable to query this type of content. $5e^{x+y}$ is developed as a first attempt to address this limitation by allowing querying for records based on mathematical equations.

\section{System Overview}
In this section we provide a general overview of the main design points of $5e^{x+y}$. A more detailed description of the system can be found in \cite{mathexplorerweb}.
\subsection{Features Extraction}
Our initial focus in the project was to define a good set of features that we could extract and index from our equations. We divided them, into two different categories. The notational features relate to the actual naming and representation of the different elements of the equation. The selected format for representing and internal processing of the equations was MathML because it allows an easier processing.

For notational features, we extracted the leave elements of the equation: variables, numbers and operator represented by the tags <mi>, <mn> and <mo> respectively. We indexed as well 2-grams and 3-grams of symbol-operator and symbol-operator-symbol.
We perform some processing of these tokens such as normalizing and decomposing of complex characters. This allowed us to partially match characters with different unicode points but similar or equal visual rendering. For instance LATIN CAPITAL LETTER A WITH RING ABOVE with code 0xc5 and ANGSTROM SIGN with code 0x212b both are rendered as \r{A} and both can be used to refer to the Angstrom measurement unit. Following the example, we indexed both the original token <mi>\r{A}</mi> and the normalized form <mi>\r{A}</mi> <mi>A</mi>. Another set of rules that we applied take into account the different types of operators that are semantically related but whose unicode representation does not provide meaningful ways to match them. We grouped related operators, for example the INTEGRAL character with code 0x222b generates the tokens <mo>{\unicodefont ∫}</mo> and INT\_OP and the CONTOUR INTEGRAL character with code 0x222e generates the tokens <mo>{\unicodefont ∮} </mo> and INT\_OP. In this way, both have a partial matching in the second token.

The second set of features relate to the structure of the equation. Two different properties of mathematical equations impose additional challenges for a MIR system. First, variables can be used indistinguishable and any character set of characters can be used to represent the same equation. $f(t) = vt + c $ and $g(x) = ac + b$ can represent the same linear equation though the context in which each variable is used follows certain conventions. For addressing this issue, we relied on pattern matching, a very powerful feature that current Computer Algebra System (CAS) provide. We integrated Mathematica into our system, to perform pattern matching on the equations and identify the occurrence of a predefined of patterns from a predefined set. This allows us to detect if the equation contained certain types of common algebraic structures. For instance, the pattern {\codefont a\_.x\^{} 2 + b\_.x + c\_.} is used to identify quadratic expression. This mechanism is very powerful since \.x represent any type of expression, not only a single term.

A second additional challenge that math equations pose is the fact that a given equality can be rewritten in multiple ways taking into account properties of the different operators like associativity, distributivity, commutativity among others. Our approach was to rely again in the features provided by Mathematica and apply the {\codefont Simplify} operations which applies a set of predefined set of rules in a equation with the goal of having a unique and concise representation.

\subsection{System architecture}
$5e^{x+y}$ is implemented in top of the Lucene/Solr framework. We expanded the default schema with  several fields, being the most important ones: {\codefont math\_notational\_field} which stores all the described tokens and n-grams, \\ {\codefont math\_structural\_field} which stores the occurrences of predefined algebraic structures and {\codefont filename} which is used to store the file/record where the particular equation was found. 
In order to extend Lucene with the new use case, our work extends some of the base classes as follows: {\codefont MultiplePatternTokenizer} is in charge of producing a stream of tokens from a single equations. This tokens are identified by applying multiple regular expressions. {\codefont UnicodeNormalizingFilter} applies the unicode normalizing transformation described above. {\codefont SynonymExpandingFilter} expands operators, with a manually computed table, with the category they belong to. Finally {\codefont StructuralPatternsTokenizer} provides the pattern matching functionality by establishing  a communication with Mathematica's {\codefont Kernel Link} API.

\section{Demonstration}
As part of the demonstration of $5e^{x+y}$, we present a demo instance of the Invenio software with the searching by mathematical equations functionality.
The user access the math searching functionality through a link in the main page. Once there, the user is presented with a classic search functionality where the user inputs his equation. He can choose between writing the query on \LaTeX or MathML. For writing the \LaTeX code, we include an online editor\cite{latex_editor} that renders the the input as the user types it. The first format is normally more familiar to users , however, the internal representation is in MathML since it allows an easier processing. The translation from \LaTeX  to MathML is done through the third party library SnuggleTex\cite{snuggletex} that have some shortcomings, so we allow users to input the query in MathML as well. When the query is ready, the user clicks the submit button and the system performs the search.  Once the system has performed the lookup and got the results, these are presented in the results interface. 
The system shows up the 20 results with highest score. Each result shows the matched equation snippet, the title of the record containing the given equation and the link to the record.

%\begin{figure}
%\center
%\includegraphics[height=3 cm]{images/figures/input_interface.png}
%\label{input_int}
%\caption{Web interface of $5e^{x+y}$: Query input screen}
%\end{figure}
%
%\begin{figure}
%\center
%\includegraphics[height=3 cm]{images/figures/results_interface.png}
%\label{results_int}
%\caption{Web interface of $5e^{x+y}$: Results screen}
%\end{figure}

\begin{figure}
\centering
\begin{minipage}{.27\textwidth}
  \includegraphics[width=0.9\linewidth]{images/figures/input_interface.png}
  \captionof{figure}{Input interface.}
  \label{fig:test1}
\end{minipage}%
\begin{minipage}{.27\textwidth}
  \includegraphics[width=0.9\linewidth]{images/figures/results_interface.png}
  \captionof{figure}{Result page.}
  \label{fig:test2}
\end{minipage}
\end{figure}

\section{Conclusions}

\section{Availability}
Both, the Invenio software and the $5e^{x+y}$ module are open source available through GPL Licence. Invenio can be downloaded from \url{http://invenio-software.org/}. $5e^{x+y}$ can be downloaded from \\ \url{https://github.com/arthoviedo/cern_math_explorer} .

%ACKNOWLEDGMENTS are optional
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{bibliography}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\end{document}